A Gaussian process model of human electrocorticographic data

### 2020-02-14 paper/main.tex
backing up progress

%%%%%%%%% STOPPED HERE

In neurosurgical applications where one wishes to infer full-brain
activity patterns, can our framework yield insights into where the
electrodes should be placed?  A basic assumption of our approach (and
of most prior ECoG work) is that electrode recordings are most
informative about the neural activity near the recording surface of
the electrode.  But if we consider that activity patterns throughout
the brain are meaningfully correlated, are there particular
implantation locations that, if present in a patient's brain, yield
especially high reconstruction accuracies throughout the rest of the
brain?  For example, one might hypothesize that brain structures that
are heavily interconnected with many other structures could be more
informative about full-brain activity patterns than comparatively
isolated structures.

To gain insights into whether particular electrode locations might be
especially informative, we first computed the average reconstruction
accuracy across all of each patient's electrodes (using the
across-patients cross validation test; black histograms in
Fig.~\ref{fig:corrmap}A and B).  We labeled each patient's electrodes
in each dataset with the average reconstruction accuracy for that
patient.  In other words, we assigned every electrode from each given
patient the same value, reflecting how well the activity patterns at
those electrodes were reconstructed on average.  Next, for each voxel
in the 4~mm$^3$ MNI brain, we computed the average value across any
electrode (from any patient) that came within 20 MNI units of that
voxel's center.  Effectively, we computed an \textit{information
  score} for each voxel, reflecting the average reconstruction
accuracy across any patients with electrodes near each voxel-- where
the averages were weighted to reflect patients who had more electrodes
implanted near that location. This yielded a single map for each
dataset, highlighting regions that are potentially promising
implantation targets in terms of providing full-brain activity
information via SuperEEG (Fig.~\ref{fig:informap}A and B).  Despite task
and patient differences across the two datasets, we nonetheless found
that the maps of the most promising implantation targets derived from
both datasets were similar (voxelwise correlation between information
scores across the two datasets: $r = 0.18, p < 10^{-10}$).  While the
correspondence between the two maps was imperfect, our finding that
there were some commonalities between the two maps lends support to
the notion that different brain areas are differently informative
about full-brain activity patterns.  We also examined the intersection
between the top 10\% most informative voxels across the two datasets
(gray areas in Fig.~\ref{fig:informap}C, networks shown in
Fig.~\ref{fig:networks}A (All)).
Supporting the notion that structures that are highly interconnected
with the rest of the brain might be especially good targets for
implantation, this intersecting set of voxels with the highest
information scores included major portions of the dorsal attention
network (e.g., inferior parietal lobule, precuneus, inferior temporal
gyrus, thalamus, and striatum) as well as some portions of the default
mode network (e.g., angular gyrus) that are highly interconnected with
a large proportion of the brain's gray matter~\citep[e.g., ][]{TomaVolk11}.

@@

We also wanted to see how the most informative regions changed by
frequency. Lower frequency oscillations, such as $\delta$, $\theta$, and $\alpha$, have
been posited to be involved in processing modes which span large swaths of
cortical regions and are thought to serve the purpose of integration
across many cortical sites ~\citep{Knya12}. Higher frequency
oscillations such as $\beta$ and $\gamma$, are more limited to topographic
areas ~\citep{Nune95}, and are modulated by slow oscillations
hierarchically ~\citep{EngeEtal10}. Using the frequency-specific reconstruction accuracies, we
followed the process outlined in Fig.~\ref{fig:informap}C, to obtain
the most informative locations for each frequency.  When we examined the intersection
between the top 10\% most informative voxels across the two datasets
for each bandpassed frequency, we found a somewhat consistent hierarchical
structure (see Fig.~\ref{fig:networks}E).

>>>>

A basic assumption of our approach (and
of most prior ECoG work) is that electrode recordings are most
informative about the neural activity near the recording surface of
the electrode.  But if we consider that activity patterns throughout
the brain are meaningfully correlated, are there particular
implantation locations that, if present in a patient's brain, yield
especially high reconstruction accuracies throughout the rest of the
brain?  For example, one might hypothesize that brain structures that
are heavily interconnected with many other structures could be more
informative about full-brain activity patterns than comparatively
isolated structures.  To test this hypothesis, we computed the average reconstruction
accuracy across all of each patient's electrodes (using our
across-patients cross validation test; black histograms in
Fig.~\ref{fig:corrmap}A and B).  We first labeled each patient's electrodes,
in each dataset, with the average reconstruction accuracy for that
patient.  In other words, we assigned every electrode from each
patient the same value, reflecting how well the activity patterns for that patient were reconstructed.  Next, for each voxel
in the 4~mm$^3$ MNI brain, we computed the average value across any
electrode (from any patient) that came within 20 MNI units of that
voxel's center.  This yielded an \textit{information
  score} for each voxel, reflecting the average reconstruction
accuracy across any patients with electrodes near each voxel-- where
the averages were weighted to reflect patients who had more electrodes
implanted near that location. We created a single map of these information scores for each
dataset, highlighting regions that are especially informative about activity in \textit{other} brain areas
(Fig.~\ref{fig:informap}A and B).  Despite task
and patient differences across the two datasets, we nonetheless found
that the maps of the most promising implantation targets derived from
both datasets were similar (voxelwise correlation between information
scores across the two datasets: $r = 0.18, p < 10^{-10}$).  While the
correspondence between the two maps was imperfect, our finding that
there were some commonalities between the two maps lends support to
the notion that different brain areas are differently informative
about full-brain activity patterns.  We also examined the intersection
between the top 10\% most informative voxels across the two datasets
(gray areas in Fig.~\ref{fig:informap}C, networks shown in
Fig.~\ref{fig:networks}A (All)).
Supporting the notion that structures that are highly interconnected
with the rest of the brain might be especially good targets for
implantation, this intersecting set of voxels with the highest
information scores included major portions of the dorsal attention
network (e.g., inferior parietal lobule, precuneus, inferior temporal
gyrus, thalamus, and striatum) as well as some portions of the default
mode network (e.g., angular gyrus) that are highly interconnected with
a large proportion of the brain's gray matter~\citep[e.g., ][]{TomaVolk11}.

@@

We also wondered whether the map of information scores might vary as a function of the spectral
components of the activity patterns under consideration.  For example, low frequency $\delta$ activity in the thalamus has been implicated in
supporting connectivity with cortex~\citep[e.g.,][]{Jone09}; $\theta$ activity
in frontal regions have been implicated in memory consolidation~\citep{NishEtal09} and integration~\citep{BackEtal16}; parietal and occipital $\alpha activity has been
implicated in attention modulation~\citep[e.g., see][]{Klim12}; $\beta$ activity in motor cortex have been implicated in motor control and processing~\citep{BrovEtal04, KoelEtal08, RubiEtal06};
and occipital $\gamma$ activity has been implicated in visual attention and visual processing~\citep{ZhouEtal16, TallEtal05, FrieEtal07}.
We computed analogous maps of information scores for each individual frequency band.  Across Datasets 1 and 2 we observed a weak, but statistically reliable, correspondence in
these maps, aside from in the $\alpha$ band ($\delta$: $r = 0.09, p < 10^{-57}$; $\theta$: $r = 0.24, p < 10^{-60}$; $\alpha$: $r = -0.03, p < 0.001$, $\beta$: $r = 0.02, p - 0.0011$; $\gamma_L$: $r = 0.1, p < 10^{-67}$; $\gamma_H$: $r = 0.03, p < 10^{-7}$; broadband: $r = 0.21, p < 10^{-297}).

%%%% STOPPED HERE...

as $\delta$, $\theta$, and $\alpha$, have
been implicated in processing modes which span large swaths of
cortical regions and are thought to serve the purpose of integration
across many cortical sites ~\citep{Knya12}. Higher frequency
oscillations such as $\beta$ and $\gamma$, are more limited to topographic
areas ~\citep{Nune95}, and are modulated by slow oscillations
hierarchically ~\citep{EngeEtal10}. Using the frequency-specific reconstruction accuracies, we
followed the process outlined in Fig.~\ref{fig:informap}C, to obtain
the most informative locations for each frequency.  When we examined the intersection
between the top 10\% most informative voxels across the two datasets
for each bandpassed frequency, we found a somewhat consistent hierarchical
structure (see Fig.~\ref{fig:networks}E).
